---
title: "Airline Industry"
author: "Parv Patni"
date: "December 22, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Mini Project to explain the factors for difference in price between an economy ticket and a premium-economy airline ticket

```{r}
getwd()
airline.df<-read.csv(paste("SixAirlinesDataV2.csv",sep=""))
head(airline.df)
```
```{r}
library(psych)
describe(airline.df)
```
##Mean Pricing for economy and premium in different airlines.
```{r}
aggregate(x = airline.df[c("PriceEconomy","PricePremium")],by=list(Airline=airline.df$Airline),FUN = mean)
```
```{r}
boxplot(airline.df$PriceEconomy~airline.df$Airline, main="Airline Vs PriceEconomy",col = c("lightgreen","lightblue","peachpuff","gray","yellow","cyan"))
boxplot(airline.df$PricePremium~airline.df$Airline, main="Airline Vs PricePremium",col = c("lightgreen","lightblue","peachpuff","gray","yellow","cyan"))
```

```{r}
boxplot(airline.df$PriceEconomy,airline.df$PricePremium,ylab="Prices", main="Economy vs Premium Pricing", col=c("yellow" , "lightgreen"), names=c("Economy","Premium"))
```
```{r}
par(mfrow=c(1,2))
with(airline.df,plot(Aircraft,PriceEconomy,col=c("peachpuff","khaki"),main="Aircraft vs Economy Pricing"))
with(airline.df,plot(Aircraft,PricePremium,col=c("peachpuff","khaki"), main="Aircraft vs Premium Pricing"))
```
```{r}
plot(airline.df$Airline, airline.df$SeatsEconomy, main="Airline vs Seats in Economy Class",col="sienna1")
plot(airline.df$Airline, airline.df$SeatsPremium, main="Airline vs Seats in Premium Class",col="sienna1")
```
```{r}
plot(airline.df$IsInternational,main = "Number of Domestic vs International Flights",col="khaki")
```


##Price variation of economy with flight duration
```{r}
plot(airline.df$FlightDuration,airline.df$PriceEconomy,
     main="Flight duration vs Economy Price",
     xlab="Flight duration",
     ylab = "Economy Price")
abline(lm(airline.df$PriceEconomy~airline.df$FlightDuration),
       col="red")
```

##Price variation of premium with flight duration

```{r}
plot(airline.df$FlightDuration,airline.df$PricePremium,
     main="Flight duration vs Premium Price",
     xlab="flight duration",
     ylab="Premium Price")
abline(lm(airline.df$PricePremium~airline.df$FlightDuration),
       col="blue")
```

Increase is gradual for both classes along with the Duration of flight. However, Rate of increase for economy is more compared to premium.

```{r}
attach(airline.df)
plot(WidthDifference,PriceRelative,main = "Analysis of width of Seats in Difference in price of class")
abline(lm(PriceRelative~WidthDifference),col="red")
```
```{r}
plot(PitchDifference,PriceRelative,main = "Analysis of Pitch of Seats in Difference in price of class")
abline(lm(PriceRelative~PitchDifference),col="red")
```
```{r}
library(corrgram)
corrgram(airline.df, order=TRUE, lower.panel=panel.shade,
         upper.panel=panel.pie, text.panel=panel.txt,
         main="Corrgram of airline dataset")
```

Null Hypothesis : There is no significant difference between the price of tickets in economy class and premium economy class
```{r}
t.test(PriceEconomy,PricePremium,var.equal=TRUE, paired=FALSE)
```
We have a p-value = 1.544e-11 which is less than 0.05, suggesting that there is a significant difference between PriceEconomy and PricePremium. Hence we reject our null hypothesis.

Regression model for showing the price differnce with respect to pitch difference and width difference

```{r}
fit <- lm(PriceRelative ~ PitchDifference + WidthDifference + FlightDuration + PercentPremiumSeats, data = airline.df)
summary(fit)
```

PriceRelative = -0.179033 + (0.059311)x(PitchDifference) + (0.118140)x(WidthDifference) + (0.021707)x(FlightDuration)

There is a significant relation of width difference, pitch difference and flight duration on difference in the price of tickets as their p-values <0.05, while PercentPremiumSeats contributes negatively with p-value>0.05